On the characteristic and diameter of planar integral point sets \footnote{ The work of the first author was partially supported by the Theoretical Physics and Mathematics Advancement Foundation “BASIS”. }

### 2025-08-24 char-4k-bound/char-4k-bound.tex
[4k3] Thank the Referees

	In 1945, Erdős gave an elegant proof~\cite{anning1945integral,erdos1945integral} that every IPS is finite.
	He considered an arbitrary IPS $M$ and chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.  Applying the same argument to the line segment $M_1M_3$, Erdös concluded that $\#M \le 4 \cdot |M_1M_2| \cdot |M_1M_3|$, where $\#M$ stands for cardinality of $M$.

	Thus, we can easily infer the lower bound for diameter of an IPS $M$:
	$$
		\operatorname{diam} M \ge \frac{\sqrt{\#M}}{2}.
	$$
	That was the first lower bound for the diameter ever;
	however, Erdős did not only estimate the diameter,
	but also established the toolset for further investigation, the core of which is a system of cofocal hyperbolas.

@@

	In~\cite{our-vmmsh-2018-translit} the constant (for $\# M \geq 4$) was improved to $0.3457$
	employing Point Packing in a Square Problem~\cite{markot2005newverified,costa2013valid}.
	In~\cite{my-pps-linear-bound-2019} the approach has been further developed, and the constant has been tightened to $\frac{5}{11}$.
	Finally, in~\cite{my-semi-general-5-4-bound-2019} for IPS $M$ in semi-general position (that is an IPS with no collinear triples) it was proved that
	\begin{equation}
		\label{eq:chebsb_lower_bound}
		\diam M \geq \left(\frac {\# M} 5 \right)^{5/4}
	\end{equation}
	(in assumption that the set has at east 4 points).
	All these lower bounds are based on the Erdős's framework --- cofocal hyperbolas.

@@

	In the present paper, we put the power of Erdős's and Kemnitz's approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
	We prove that for a particular case of characteristic $4k+1$ and $4k+2$
	the bound~\eqref{eq:chebsb_lower_bound} can be improved to
	\begin{equation}
		\operatorname{diam} M \ge \left( \frac{25}{36} n \right)^{5/4}
		.
	\end{equation}

	In Section 2, we give all the required notions and known results. In Section 3, we discuss some examples of integral point sets in order to demonstrate that none of the classes $4k+1$, $4k+2$, $4k+3$ is too exotic nor pathological. Section 4 is devoted to the connection between Erdős curves and the characteristic. In Section 5, we prove some auxiliary results, and we proceed to the main one in Section 6.

@@

In this section, we provide rigorous definitions and list some known results.

@@

\begin{definition}
	\label{def:char}
	The characteristic of a planar IPS $M$ is a squarefree number $q$
	such that the area of any triangle $M_1M_2M_3$, $\{M_1,M_2,M_3\}\subset M$,
	is commensurable with $\sqrt{q}$.
	Notation: $\chr M = q$.
\end{definition}
The characteristic of an IPS always exists and is unique.

@@

\begin{definition}
	Let $M_1 M_2$ be an edge of $M\in \mathfrak{M}$.
	For an integer $n$, $|n| < |M_1 M_2|$,
	we will call the set of points
	\begin{equation}
		\{M_0 : |M_0 M_1| - |M_0 M_2| = n\}
	\end{equation}
	the \emph{$n$-th Erdős curve}.
\end{definition}

Obviously, the 0-th Erdős curve is the perpendicular bisector of $M_1 M_2$,
and all the other Erdős curves are branches of cofocal hyperbolas.
On Figure~\ref{fig:Erdos_curves_numbered}, the Erdős curves are shown for an edge of length 3.

\begin{figure}
\center{\includegraphics[width=0.95\linewidth]{Erdos_curves_numbered.png}}
\caption{Erdős curves}
\label{fig:Erdos_curves_numbered}
\end{figure}

Thus, an edge $M_1 M_2$ generates $2|M_1 M_2| - 1$ Erdős curves.
For the sake of brevity, Erdős curves with odd numbers are named odd Erdős curves,
and the ones with even numbers are named even Erdős curves.

@@

\begin{lemma}
	\cite[Theorem 5.9]{my-pps-linear-bound-2019}
	\label{lem:no_distance_one}
	Each set $M\in\mathfrak{M}_n$
	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds,
	consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line,
	and one point out of the line, on the perpendicular bisector of line segment $M_1 M_2$.
\end{lemma}

@@

For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018-translit}:
$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
which means that each abscissa is multiplied by $1/q$
and each ordinate is multiplied by $\sqrt{p}/q$,  i.e.
\begin{equation}
	\label{eq:char_lattice}
	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
	=
	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}
	,
\end{equation}
where $q$ is the characteristic of the IPS.
Such notation is made possible by the Grid Theorem.
Note that we use the asterisk instead of the centered dot to emphasize
that we do not mean the usual coordinatewise multiplication;
instead, the denominator is appended to both coordinates,
but the square root multiplier is appended only to the second one.

@@

%4
\section{Erdős Curves and Characteristic}

@@

\begin{theorem}[The Weeding Theorem I]
	\label{thm:weeding_I}
	Let $M = \{M_1, M_2, M_3\} \in \mathfrak{M}_3$
	be an IPS with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$,
	ant let $M_3$ belong to the $n$-th Erdős curve of the edge $M_1 M_2$.
	Then, $n$ and the length of $M_1M_2$ have the same parity.
\end{theorem}

\begin{proof}
	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and $n = |M_1 M_3| - |M_2 M_3|$.
	If $m$ is an odd number, then by Lemma~\ref{lem:pre_weeding_I}, the difference $n = |M_1 M_3| - |M_2 M_3|$ equals $m - s$, where $s$ is an even number, is also an odd number.
	Let $m$ be an even number; then $n = |M_1 M_3| - |M_2 M_3| = m - s$ is an even number, as a difference of two even numbers.
\end{proof}

@@

\begin{theorem}[The Weeding Theorem II]
	\label{thm:weeding_II}
	Let an IPS $M=\{M_1, M_2, M_3\}\in{\mathfrak{M}_{3}}$ have an even edge length $|M_1 M_2|$, and let point $M_3$ lie on an odd Erdős curve of edge $M_1 M_2$.
	Then $\operatorname{char} M=8k+7$, $k\in \mathbb{N}_{0}$.
\end{theorem}

@@

\begin{lemma}
	\label{lem:no_distance_2}
	Every set $M \in \overline{\mathfrak{M}}'_n$ with an edge $|M_1M_2|=2$ has the cardinality $n=3$.
\end{lemma}

\begin{proof}
	Let $M \in \overline{\mathfrak{M}}'_n$ and $M_1, M_2 \in M$.
	Then all points in the set $M$ lie on $cr(M_1, M_2)$.
	Otherwise, according to Lemma~\ref{lem:pre_weeding_II}, $\operatorname{char}{M}={8k+7}$, $k\in \mathbb{N}_0$.

	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}}'_4$.
	Employ the Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
	Then $M_3$ and $M_4$ lie on the perpendicular bisector to $M_1M_2$.
	Since the distance $|M_3M_4|$ is an integer, the area of triangle $M_1M_3M_4$ is a rational number.
	By Definition~\ref{def:char}, this means that $\operatorname{char}{M}={1}$.
	Then by the Grid Theorem we have $M_3(0;t/4)$. Without loss of generality, let us assume that $t\in\mathbb{N}$. Let $|M_1M_3|=s$. Applying the Pythagorean theorem to triangle $OM_1M_3$, we get
	$$
	1+\frac{t^2}{16}=s^2,
	$$
	or equivalently,
	\begin{equation}
	\label{eq: 42}
	16s^2-t^2=16.
	\end{equation}
	Now, we find the solutions to equation~\eqref{eq: 42} in positive integers.
	To do this, we introduce the substitution: $t=4k$, $k\in\mathbb{N}$.
	Then equation~\eqref{eq: 42} takes the form:
	\begin{equation}
	\label{eq: 43}
	s^2-k^2=1.
	\end{equation}
	Break down the left-hand side of equation~\eqref{eq: 43} into factors:
	\begin{equation}
	\label{eq: 44}
	(s-k)(s+k)=1.
	\end{equation}
	Since $s,k\in\mathbb{N}$, the number 1 can be represented as the product of two integers in two ways: $1\cdot1$ and $-1\cdot(-1)$. In the first case, $(s-k)=1$ and $(s+k)=1$, which gives $s=1,k=0$, which contradicts that $k\in\mathbb{N}$.
	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1$, $k=0$, which also contradicts that $s,k\in\mathbb{N}$.

@@

\begin{corollary}
	\label{cor:sqrt_5_3}
	In a triangle $ABC$ with sides $3 \leq a \leq b \leq c$, where ${A, B, C} \subset M \in \overline{\mathfrak{M}}'_4$, the smallest height is at least $(2a-1)^{1/2}$.
\end{corollary}

\begin{proof}
	Indeed, for $a\geq 3$ the inequality
	$$
		(2a-1)^{1/2} \geq \frac{\sqrt{5}}{\sqrt{3}} a^{1/2}
	$$
	holds, and
	the difference between the left and right-hand sides increases with increasing $a$.
\end{proof}

@@

The following lemma is a classical Erdős-style intersection-enumeration one,
empowered by our Weeding Theorem I.

\begin{lemma}
	\label{lem:intersection_enumeration}
	Let ${M_1, M_2, M_3, M_4} \subset M \in \overline{\mathfrak{M}}'_n$ (points $M_2$ and $M_3$ may coincide, while the others are distinct), where $n \ge 4$. Then $\# M \le |M_1 M_2| \cdot |M_3 M_4| - 2$.
\end{lemma}

@@

\begin{equation}
	\begin{array}{rcl}
		\label{d}
		\overline{d}(2,3)&=& 1,{}\\
		\overline{d}(2,4)&=& 4,{} \\
		\overline{d}(2,5)&=&  \overline{d}(2,6)= 8,{} \\
		\overline{d}(2,7)&=& 33,{} \\
		\overline{d}(2,8)&=&  \overline{d}(2,9)= 56,{} \\
		\overline{d}(2,10)&=& ... =\overline{d}(2,12)= 105,{} \\
		\overline{d}(2,13)&=&  \overline{d}(2,14)= 532,{}\\
		\overline{d}(2,15)&=&  ... = \overline{d}(2,18) = 735,{} \\
		\overline{d}(2,19)&=& ... = \overline{d}(2,24) = 1995,{} \\
		\overline{d}(2,25)&=&  ... = \overline{d}(2,27)= 9555,{} \\
		\overline{d}(2,28)&=& 10672,{} \\
		\overline{d}(2,29)&=&  ... = \overline{d}(2,36) = 13975,{} \\
		\overline{d}(2,37) &>& 20\ 000.
	\end{array}
\end{equation}

Now we are finally ready to proove our main result.
It improves the bound~\eqref{eq:chebsb_lower_bound}
in our special case.

\begin{theorem}
	\label{thm:main_result}
	Let $M \in \overline{\mathfrak{M}}'_n$, i.e., $M$ is a set in semi-general position with characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
	Then, for every integer $n \geq 3$, the inequality
	\begin{equation}
		\operatorname{diam} M \geq \left( \frac{25n}{36} \right)^{5/4}
	\end{equation}
	holds.
\end{theorem}

@@

	Let us consider $M \in \overline{\mathfrak{M}}'_n$, $n \geq 4$, $\operatorname{diam} M = p$.

@@

	If $n\geq 4$ and $p < 10$, then the claim of the theorem follows immediately from~\eqref{d}.
	So now we have to consider $m > \frac{6}{5}p^{2/5} > 3$.
	Then for any points $A, B \in M \setminus {M_1}$, we have $|AB| > \frac{6}{5}p^{2/5}$.
	By Corollary~\ref{cor:sqrt_5_3} and Lemma~\ref{lem:smurov_min_height},
	no three points from $M \setminus {M_1}$ lie in a strip of width
	\begin{equation}
		\frac{\sqrt{5}}{\sqrt{3}} \cdot \sqrt{\frac{6p^{2/5}}{5}} = \sqrt{2} \cdot p^{1/5}
		.
	\end{equation}

@@

	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the theorem is true.
	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter is also true.
	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10 655$.
	This estimate is weaker than the available numerical results. Therefore, from now on, we can assume that $n \geq 7^4$.

@@

\section{Acknowledgements}
The authors thank Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
Dr. A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
Dr.~A.S.~Usachev for proofreading,
and E.A. Momot for the ``SciLexic'' project that helped the authors with some word usage issues.
Also the authors would like to owe a special thanks to the anonymous referees
for their helpful comments and suggestions on the paper.

>>>>

	In 1945, Erdös gave an elegant proof~\cite{anning1945integral,erdos1945integral} that every IPS is finite.
	He considered an arbitrary IPS $M$ and chose a non-collinear triple $\{M_1, M_2, M_3\} \in M$ so that any other point $M_0 \in M$ lies on either the straight line $M_1M_2$, the perpendicular bisector to the segment $M_1M_2$, or one of $|M_1M_2|-1$ hyperbolas, where $|M_1M_2|$ stands for length of line segment $M_1M_2$.  Applying the same argument to the line segment $M_1M_3$, Erdös concluded that $\#M \le 4 \cdot |M_1M_2| \cdot |M_1M_3|$, where $\#M$ stands for cardinality of $M$.

	Thus, we can easily infer the lower bound for diameter of an IPS $M$:
	$$
		\operatorname{diam} M \ge \frac{\sqrt{\#M}}{2}.
	$$
	That was the first lower bound for the diameter ever;
	however, Erdös did not only estimate the diameter,
	but also established the toolset for further investigation, the core of which is a system of cofocal hyperbolas.

@@

	In~\cite{our-vmmsh-2018-translit} the constant (for $\# M \geq 4$) was improved to $0.3457$
	employing Point Packing in a Square Problem~\cite{markot2005newverified,costa2013valid}.
	In~\cite{my-pps-linear-bound-2019} the approach has been further developed, and the constant has been tightened to $\frac{5}{11}$.
	Finally, in~\cite{my-semi-general-5-4-bound-2019} for IPS $M$ in semi-general position (that is an IPS with no collinear triples) it was proved that
	\begin{equation}
		\label{eq:chebsb_lower_bound}
		\diam M \geq \left(\frac {\# M} 5 \right)^{5/4}
	\end{equation}
	(in assumption that the set has at east 4 points).
	All these lower bounds are based on the Erdos's framework --- cofocal hyperbolas.

@@

	In the present paper, we put the power of Erdös's and Kemnitz's approaches together. Due to squarefree nature of characteristic, it can be of the form $4k+1$, $4k+2$ or $4k+3$, where $k$ is integer.
	We prove that for a particular case of characteristic $4k+1$ and $4k+2$
	the bound~\eqref{eq:chebsb_lower_bound} can be improved to
	\begin{equation}
		\operatorname{diam} M \ge \left( \frac{25}{36} n \right)^{5/4}
		.
	\end{equation}

	In Section 2, we give all the required notions and known results. In Section 3, we discuss some examples of integral point sets in order to demonstrate that none of the classes $4k+1$, $4k+2$, $4k+3$ is too exotic nor pathological. Section 4 is devoted to the connection between Erdös curves and characteristic. In Section 5, we prove some auxiliary results, and we proceed to the main one in Section 6.

@@

In this Section, we provide rigorous definitions and list some known results.

@@

\begin{definition}
	\label{def:char}
	The characteristic of a planar IPS $M$ is a squarefree number $q$
	such that the area of any triangle $M_1M_2M_3$, $\{M_1,M_2,M_3\}\subset M$,
	is commensurable with $\sqrt{q}$.
	Notation: $\chr M = q$.
\end{definition}

@@

\begin{definition}
	Let $M_1 M_2$ be an edge of $M\in \mathfrak{M}$.
	For $N\in\mathbb N_\pm$, $|n| < |M_1 M_2|$
	we will say that the set of points
	\begin{equation}
		\{M_0 : |M_0 M_1| - |M_0 M_2| = n\}
	\end{equation}
	is called the \emph{$n$-th Erdos curve}.
\end{definition}

Obviously, the 0-th Erdos curve is the perpendicular bisector of $M_1 M_2$,
and all the other Erdos curves are branches of cofocal hyperbolas.
On Figure~\ref{fig:Erdos_curves_numbered}, the Erdos curves are shown for an edge of length 3.

\begin{figure}
\center{\includegraphics[width=0.95\linewidth]{Erdos_curves_numbered.png}}
\caption{Erdos curves}
\label{fig:Erdos_curves_numbered}
\end{figure}

Thus, an edge $M_1 M_2$ generates $2|M_1 M_2| - 1$ Erdos curves.
For the sake of brevity, Erdos curves with odd numbers are named odd Erdos curves,
and the ones with even numbers are named even Erdos curves.

@@

\begin{lemma}
	\cite[Theorem 3.10]{my-pps-linear-bound-2019}
	\label{lem:no_distance_one}
	Each set $M\in\mathfrak{M}_n$
	such that for some $M_1,M_2 \in M$ equality $|M_1 M_2|=1$ holds,
	consists of $n-1$ points, including $M_1$ and $M_2$, on a straight line,
	and one point out of the line, on the perpendicular bisector of line segment $M_1 M_2$.
\end{lemma}

@@

For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018-translit}:
$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
which means that each abscissa is multiplied by $1/q$
and each ordinate is multiplied by $\sqrt{p}/q$,  i.e.
\begin{equation}
	\label{eq:char_lattice}
	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
	=
	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}
	,
\end{equation}
where $q$ is the characteristic of the IPS.
Such notation is made possible by the Grid Theorem.

@@

%4
\section{Erdös Curves and Characteristic}

@@

\begin{theorem}[The Weeding Theorem I]
	\label{thm:weeding_I}
	Let $M = \{M_1, M_2, M_3\} \in \mathfrak{M}_3$ be an IPS of semi-general position with a characteristic different from $4k+3$, $k \in \mathbb{N}_0$. Then, the length of the edge $M_1 M_2$ and its $n$-th Erdős curve, which contains the point $M_3$, have the same parity.
\end{theorem}

\begin{proof}
	Let $|M_1 M_2| = m$, $m \in \mathbb{N}$, and let the $n$-th Erdős curve satisfy $|M_1 M_3| - |M_2 M_3| = n$.
	If $m$ is an odd number, then by Lemma~\ref{lem:pre_weeding_I}, the difference $n = |M_1 M_3| - |M_2 M_3|$ equals $m - s$, where $s$ is an even number, is also an odd number.
	Let $m$ be an even number; then $n = |M_1 M_3| - |M_2 M_3| = m - s$ is an even number, as a difference of two even numbers.
\end{proof}

@@

\begin{theorem}[The Weeding Theorem II]
	\label{thm:weeding_II}
	Let an IPS in semi-general position $M=\{M_1, M_2, M_3\}\in{\mathfrak{M}_{3}}$ have an even edge length $|M_1 M_2|$, and let point $M_3$ lie on an odd Erdős curve of edge $M_1 M_2$.
	Then the set $M$ has characteristic $p=8k+7$, $k\in \mathbb{N}_{0}$.
\end{theorem}

@@

\begin{lemma}
	\label{lem:no_distance_2}
	Every set $M \in \overline{\mathfrak{M}_n}'$ with an edge $|M_1M_2|=2$ has a cardinality $n=3$.
\end{lemma}

\begin{proof}
	Let $M \in \overline{\mathfrak{M}_n}'$ and $M_1, M_2 \in M$.
	Then all points in the set $M$ lie on $cr(M_1, M_2)$.
	Otherwise, according to Lemma~\ref{lem:pre_weeding_II}, $\operatorname{char}{M}={8k+7}$, $k\in \mathbb{N}_0$.

	Suppose, to the contrary, that $M={M_1, M_2, M_3, M_4} \in \overline{\mathfrak{M}_4}'$.
	Employ the Grid Theorem and set $M_1(-1;0), M_2(1;0), O(0;0)$.
	Then $M_3$ and $M_4$ lie on the perpendicular bisector to $M_1M_2$.
	Since the distance $|M_3M_4|$ is an integer, the area of triangle $M_1M_3M_4$ is a rational number.
	By Definition~\ref{def:char}, this means that $\operatorname{char}{M}={1}$.
	Then by the Grid Theorem we have $M_3(0;t/4)$. Without loss of generality, let us assume that $t\in\mathbb{N}$. Let $|M_1M_3|=s$. Applying the Pythagorean theorem to triangle $OM_1M_3$, we get
	$$
	1+\frac{t^2}{16}=s^2,
	$$
	or equivalently,
	\begin{equation}
	\label{eq: 42}
	16s^2-t^2=16.
	\end{equation}
	Now, we find the solutions to equation~\eqref{eq: 42} in positive integers.
	To do this, we introduce the substitution: $t=4k$, $k\in\mathbb{N}$.
	Then equation~\eqref{eq: 42} takes the form:
	\begin{equation}
	\label{eq: 43}
	s^2-k^2=1.
	\end{equation}
	Break down the left-hand side of equation~\eqref{eq: 43} into factors:
	\begin{equation}
	\label{eq: 44}
	(s-k)(s+k)=1.
	\end{equation}
	Since $s,k\in\mathbb{N}$, the number 1 can be represented as the product of two integers in two ways: $1\cdot1$ and $-1\cdot(-1)$. In the first case, $(s-k)=1$ and $(s+k)=1$, which gives $s=1,k=0$, which contradicts that $k\in\mathbb{N}$.
	In the second case, $(s-k)=-1$ and $(s+k)=-1$, which gives $s=-1$, $k=0$, which also contradicts that $s,k\in\mathbb{N}$.

@@

\begin{corollary}
	\label{cor:sqrt_5_3}
	For the height of a triangle $ABC$, where ${A, B, C} \subset M \in \overline{\mathfrak{M}_4}'$, with sides $3 \leq a \leq b \leq c$, the following estimate holds:
	$$
		(2a-1)^{1/2} \geq \frac{\sqrt{5}}{\sqrt{3}} a^{1/2},
	$$
	where the difference between the left and right-hand sides increases with increasing $a$.
\end{corollary}

@@

The following lemma is a classical Erdos-style intersection-enumeration one,
empowered by our Weeding Theorem I.

\begin{lemma}
	\label{lem:intersection_enumeration}
	Let ${M_1, M_2, M_3, M_4} \subset M \in \overline{\mathfrak{M}_n}$ (points $M_2$ and $M_3$ may coincide, while the others are distinct), where $n \ge 4$. Then $\# M \le |M_1 M_2| \cdot |M_3 M_4| - 2$.
\end{lemma}

@@

\begin{eqnarray}
\label{d}
\overline{d}(2,3)&=& 1,{}\nonumber\\
\overline{d}(2,4)&=& 4,{}\nonumber\\
\overline{d}(2,5)&=&  \overline{d}(2,6)= 8,{}\nonumber\\
\overline{d}(2,7)&=& 33,{}\nonumber\\
\overline{d}(2,8)&=&  \overline{d}(2,9)= 56,{}\nonumber\\
\overline{d}(2,10)&=& ... =\overline{d}(2,12)= 105,{}\nonumber\\
\overline{d}(2,13)&=&  \overline{d}(2,14)= 532,{}\nonumber\\
\overline{d}(2,15)&=&  ... = \overline{d}(2,18) = 735,{}\nonumber\\
\overline{d}(2,19)&=& ... = \overline{d}(2,24) = 1995,{}\nonumber\\
\overline{d}(2,25)&=&  ... = \overline{d}(2,27)= 9555,{}\nonumber\\
\overline{d}(2,28)&=& 10672,{}\nonumber\\
\overline{d}(2,29)&=&  ... = \overline{d}(2,36) = 13975,{}\nonumber\\
\overline{d}(2,37) &>& 20\ 000.
\end{eqnarray}

Now we are finally ready to proof our main result.
It improves the bound~\eqref{eq:chebsb_lower_bound}
in our special case.

\begin{theorem}
	\label{thm:main_result}
	Let $M \in \overline{\mathfrak{M}}_n'$, i.e., $M$ is a set in semi-general position with characteristic different from $4k+3$, $k \in \mathbb{N}_0$.
	Then, for every integer $n \geq 3$, the inequality
	\begin{equation}
		\operatorname{diam} M \geq \left( \frac{25n}{36} \right)^{5/4}
	\end{equation}
	holds.
\end{theorem}

@@

	Let us consider $M \in \overline{\mathfrak{M}_n}'$, $n \geq 4$, $\operatorname{diam} M = p$.

@@

	Now we have to consider $m > \frac{6}{5}p^{2/5}$.
	Then for any points $A, B \in M \setminus {M_1}$, we have $|AB| > \frac{6}{5}p^{2/5}$.
	By Corollary~\ref{cor:sqrt_5_3} and Lemma~\ref{lem:smurov_min_height},
	no three points from $M \setminus {M_1}$ lie in a strip of width
	\begin{equation}
		\frac{\sqrt{5}}{\sqrt{3}} \cdot \sqrt{\frac{6p^{2/5}}{5}} = \sqrt{2} \cdot p^{1/5}
		.
	\end{equation}

@@

	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the theorem is true.
	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter is also true.
	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10,655$.
	This estimate is weaker than the available numerical results. Therefore, from now on, we can assume that $n \geq 7^4$.

@@

\section{Acknowledgements}
The author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
Dr. A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
Dr.~A.S.~Usachev for proofreading,
and E.A. Momot for the ``SciLexic'' project that helped the author with some word usage issues.
Also the author would like to owe a special thanks to the anonymous referees
for their helpful comments and suggestions on the paper.
